---
title: "Modeling trends of COVID-19 confirmed cases using linear regression!"
author: ''
date: '2021-04-02'
tags:
- R Markdown
- covid
categories: R
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```


##  Data collection and processing

First, we collect the covid-19 data from the website https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv. And do some data cleaning using R to obtain a cleaned data frame.

```{r}
library(readr)
library(dplyr)
library(tidyverse)
library(tidyr)
```



```{r}
confirmedraw <- read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv", head = TRUE,stringsAsFactors = FALSE)

regions <- sort(unique(confirmedraw$Country.Region))
n <- length(regions)
df <- c()
for(i in 1:n) {
  d <- confirmedraw[confirmedraw$Country.Region == regions[i],]
  df <- cbind(df, apply(d[,-c(1:4)], 2, sum))
}
row.names(df) <- 1:nrow(df)
df <- data.frame(df)
colnames(df) <- regions
dates <- seq(as.Date("2020-01-22"),
    as.Date(gsub("X","",colnames(confirmedraw)[ncol(confirmedraw)]),"%m.%d.%y"),1)

df <- mutate(df, 
             date = dates,
             total = rowSums(df[,-ncol(df)]))
```

## Data visualizations

```{r}
df %>% filter( as.Date("2020-01-22") <= date & date <= as.Date("2021-03-31")) %>%
ggplot(aes(date, total)) + geom_line() + ylab("Cumulative confirmed cases") +
  theme_bw()
```


Figure 1 shows the global trend of covid-19, it can be seen clearly the world faced with serious situation that the confirmed cases of covid-19 is still increasing quickly from the 2020-01 to 2021-03. 


```{r,fig.width=7.6}
df2 <- df[,c("date","China","US","Canada","Germany")]

df3 <- gather(df2, "Region", "count", -date)

df3 %>% ggplot( aes(date, count, color = Region))  + geom_line() + 
     ylab("Cumulative confirmed cases") +
     theme_bw()

```

Figure 2 shows the trends of covid-19 in the 4 countres "China","US","Canada","Germany" respecitvely, clearly, it can be seen the cases in US are most and situation is worst that the increasing speed and cases are both highest. For China, the line is almost horizonal now while the countries Canada and Germany are in the middle.


```{r}
df3 %>% ggplot( aes(date, count, color = Region))  + geom_line() + 
     ylab("Log scale of cumulative confirmed cases") +
     theme_bw() + scale_y_log10()
```

Figure 3 shows the trends of covid-19 in log scales for the above 4 countries, clearly, it shows the speed of the increasing of new confirmed cases, that China has almost with a speed 0 compared the other 3 countries, it indicates the convid-19 in China is controlled very well but in other countries, the situations are very seriously. 


```{r}
df3 %>% filter(as.Date("2020-04-01") <= date) %>% 
  ggplot( aes(date, count, color = Region))  + geom_line() + 
     ylab("Log scale of cumulative confirmed cases") +
     theme_bw() + scale_y_log10() +  geom_smooth(method = "lm", fill = NA) 
```

Figure 4  shows that linear trend models can be used to modeling the trends of the covid-19 in log-scales for main countries in the world. And clearly, China has a slope close to 0, while the US trend fits the linear model the best while the other two countries fit little worse but overall, the linear models capture the trends of increasing of new confirmed cases for these countries.


In conclusion, this study finds that for different countries in the world, the linear model could be used to model the  trends of the covid-19 in log-scales well overall, this might be very helpful to make predictions of the new confirmed cases in future, and the governments could make better decisions in dealing with the cases such as preparing enough masks, hospital beds and other resources.






